Radiance - A Web Application Environment

### 2017-03-15 els2017-radiance/paper.tex
Style/grammar corrections.

\begin{abstract}
  Radiance\cite{radiance} is a set of libraries that provide an environment for web applications. By putting its focus on running multiple web services or applications within the same environment, it has developed a features that set it apart from traditional web frameworks. In particular, it shows a new approach to the way different framework features are provided, and how the routing to content is performed. These differences allow applications written against Radiance to transparently share common resources such as the HTTP server, user accounts, the authentication, and so forth with each other.
\end{abstract}

@@

\section{Introduction}
As the internet evolved, served websites began to evolve and became more and more dynamic. Content is no longer a set of static webpages, but rather automatically updated, or even individually composed for the specific user viewing the website. Creating such dynamic websites requires a lot of extra work, much of which is usually done in much the same way for every website. The requested content needs to be retrieved somehow, transformed as needed, and finally assembled into deliverable HTML. \\

@@

For our blog to share as many resources as it can with potential third-party applications residing in the same installation, it needs to rely on the framework to provide all of our required features, except perhaps for the template engine. Additionally, it needs to have some system that divides up the URL namespace between applications. As the application writer, we cannot have any presumptions about what the final setup will look like-- what kinds of applications will be deployed together, and how the public URLs should resolve. \\

@@

\subsection{Interfaces}
Interfaces represent a form of contract. They allow you to define the signatures of constructs exposed from a package, such as functions, macros, classes, and so forth. These function signatures are accompanied by a documentation that specifies their public behaviour. Through this, Radiance can provide several ``standard interfaces'' that specify the behaviour of many features that are commonly needed in web applications, without actually having to implement them. \\

In order to make this more concrete, let's look at an interface that might provide some form of caching mechanism.

@@

Radiance extends ASDF's dependency resolution mechanism to allow for this kind of declaration. When the \\\code{blog-service} system is now loaded, Radiance will notice the interface dependency and resolve it to an actual system that implements the desired cache functionality. This resolution is defined in the currently active environment, and can thus be configured by the administrator of the Radiance installation. \\

@@

Using direct overwriting of definitions means that all applications must use the same implementation. However, usually this is the intended effect, as we want to maximise the sharing between them. If an implementation should have special needs, it can always bypass the interfaces and make direct use of whatever it might depend on. This interfaces approach does bring some benefits, however. For one, it allows the implementations to be as efficient as possible, as there is no intermediate layer of redirection. \\

Radiance provides standard interface definitions for all of the components we require for the blog application, and more. A full list of the interfaces and their descriptions is available in the Radiance documentation\footnote{\url{https://github.com/Shirakumo/radiance/\#interface}}. Thus, Radiance can be used like a macro-framework, but does not load on any features unless specifically required. Additionally, any of the implementations can be exchanged by a system administrator for ones that more closely match their requirements, without having to change anything about the application code.\\

@@

The translation between the two namespaces is the responsibility of the routing system, which is illustrated in \hyperref[requestcycle]{Figure 1}. \\

@@

When a request is dispatched by Radiance, it first parses the request URL into an object presentation that makes it easier to modify. It then sends it through the routing system's mapping functions, which turn this external URL into an internal one. The request is then dispatched to the according application that ``owns'' the URL in the internal representation. \\

@@

With the mapping alone the case is not yet solved, though. When a page is emitted that contains links, the links too must be translated, but in the opposite way. Unfortunately, because routes can be arbitrarily complex, it is not possible for the system to figure out a reversal route automatically. The logic of a reversal route will usually be much the same as it was for the mapping route and should thus not be a problem to figure out, though. \\

@@

The environment itself simply dictates a directory that contains all of these configuration files. Each application in the system automatically receives its own directory in which it can store both configuration, and temporary data files. By simply switching out this environment directory, the system can then be loaded into completely different configurations, for example allowing you to easily keep ``development'' and ``production'' setups. \\

@@

.. or instead use the programmatical way, like so:

@@

\section{Further Work}
Currently, while it is possible to dynamically load applications, it is not possible to dynamically unload them. Doing so is troublesome, as an application's system might potentially modify any other part of the Lisp process. However, if there is a constraint on what kind of changes can be rolled back, it should be possible to provide a usable form of this kind of feature. \\

@@

\section{Acknowledgements}
Thanks to Joram Shrijver, Till Ehrengruber, and Fernando Borretti for feedback and suggestions on the paper.

>>>>

\begin{abstract}
  Radiance\cite{radiance} is a set of libraries that provide an environment for web applications. By putting its focus on running multiple web services or applications within the same environment, it has developed features that set it apart from traditional web frameworks. In particular, it shows a new approach to the way different framework features are provided, and how the routing to content is performed. These differences allow applications written against Radiance to transparently share common resources such as the HTTP server, user accounts, authentication, and so forth with each other.
\end{abstract}

@@

\section{Introduction}
As the internet evolved, websites began to evolve and became more and more dynamic. Content is no longer a set of static webpages, but rather automatically updated, or even individually composed for the specific user viewing the website. Creating such dynamic websites requires a lot of extra work, much of which is usually done in much the same way for every website. The requested content needs to be retrieved somehow, transformed as needed, and finally assembled into deliverable HTML. \\

@@

For our blog to share as many resources as it can with potential third-party applications residing in the same installation, it needs to rely on the framework to provide all of our required features, except perhaps for the template engine. Additionally, it needs to have some system that divides up the URL namespace between applications. As the application writer, we cannot have any presumptions about what the final setup will look like --- what kinds of applications will be deployed together, and how the public URLs should resolve. \\

@@

\subsection{Interfaces}
Interfaces represent a form of contract. They allow you to define the signatures of constructs exposed from a package, such as functions, macros, classes, and so forth. These function signatures are accompanied by documentation that specifies their public behaviour. Through this, Radiance can provide several ``standard interfaces'' that specify the behaviour of many features that are commonly needed in web applications, without actually having to implement them. \\

In order to make this more concrete, let's look at an interface that might provide some sort of caching mechanism.

@@

Radiance extends ASDF's dependency resolution mechanism to allow for this kind of declaration. When the \\\code{blog-service} system is now loaded, Radiance will notice the interface dependency and resolve it to an actual system that implements the desired caching functionality. This resolution is defined in the currently active environment, and can thus be configured by the administrator of the Radiance installation. \\

@@

Using direct overwriting of definitions means that all applications must use the same implementation. However, usually this is the intended effect, as we want to maximise the sharing between them. If an implementation should have special needs, it can always bypass the interfaces and make direct use of whatever it might depend on. This approach with interfaces does bring some benefits, however. For one, it allows the implementations to be as efficient as possible, as there is no intermediate layer of redirection. \\

Radiance provides standard interface definitions for all of the components we require for the blog application, and more. A full list of the interfaces and their descriptions is available in the Radiance documentation\footnote{\url{https://github.com/Shirakumo/radiance/\#interface}}. Thus, Radiance can be used like a macro-framework, but does not load any features unless specifically required. Additionally, any of the implementations can be exchanged by a system administrator for ones that more closely match their requirements, without having to change anything about the application code.\\

@@

The translation between the two namespaces is the responsibility of the routing system. The way it is connected into the life-cycle of a request is illustrated in \hyperref[requestcycle]{Figure 1}. \\

@@

When a request is dispatched by Radiance, it first parses the request URL into an object presentation that makes it easier to modify. It then sends it through the routing system's mapping functions, which turn this external URL into an internal one. The request is then dispatched to the application that ``owns'' the URL in the internal representation. \\

@@

With the mapping alone the case is not yet solved, though. When a page is emitted that contains links, the links must be translated as well, but in the opposite way. Unfortunately, because routes can be arbitrarily complex, it is not possible for the system to figure out a reversal route automatically. The logic of a reversal route will usually be much the same as it was for the mapping route and should thus not be a problem to figure out, though. \\

@@

The environment itself simply dictates a directory that contains all of these configuration files. Each application in the system automatically receives its own directory in which it can store both configuration and temporary data files. By simply switching out this environment directory, the system can then be loaded into completely different configurations, for example allowing you to easily keep ``development'' and ``production'' setups. \\

@@

\dots or instead use the programmatical way, like so:

@@

\section{Further Work}
Currently, while it is possible to dynamically load applications, it is not possible to dynamically unload them. Doing so is troublesome, as an application's system might potentially modify any other part of the Lisp process. However, if a constraint on what kind of changes can be rolled back is introduced, it should be possible to provide a usable form of this kind of feature. \\

@@

\section{Acknowledgements}
Thanks to Joram Schrijver, Till Ehrengruber, and Fernando Borretti for feedback and suggestions on the paper.
